---
title: "key_test_short"
author: "Allison Horst"
date: "December 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages}

library(tidyverse)
library(lubridate)
library(janitor)
library(gganimate)
library(ggimage)
library(corrplot)

```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}

env_var <- read_csv("world_env_vars.csv") %>% 
  na.omit

cor_df <- cor(env_var[2:28])
corrplot(cor_df, 
         type = "upper", 
         method = "ellipse", 
         order = "hclust", 
         tl.col = "black", 
         tl.srt = 45, 
         tl.cex = 0.5) # also try method = number, method = color

```

# PART 2. Binary Logistic Regression

a. Download and load dataset 'DonnerParty.csv' and explore (survival for Donner Party members). We will explore odds of survival based on age and sex. In this dataset, for Outcome variable: 1 = survived, 0 = died. 

b. Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We will use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable
```{r donner_data}

DonnerTable <- read_csv("DonnerTable.csv")
```

b. Binomial logistic regression
```{r blr}

donner_blr <- glm(Outcome ~ Sex + Age, family = binomial, data = DonnerTable)
donner_blr
summary(donner_blr)

```


c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

f_5 <- data.frame(Sex = "Female", Age = 5)

# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f_5_logodds <- predict(donner_blr, newdata = f_5, type = "link") # type = "link" reports the log-odds (the original )

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

f_5_odds <- exp(f_5_logodds) # Calculate odds of survival for a 5 year old female (odds = 4.24)

# How would we manually find the probability of survival for a 5-year old female? (p/(1-p) = 4.24, so p = 4.24/5.24 = 0.809, or 80.9% probability of survival)

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f_5_prob <- predict(donner_blr, newdata = f_5, type = "response")
f_5_prob # Matches (p = 0.809)

```


d. What is the predicted probability of survival for a 25 year-old male? 

```{r predict_m_25}

m_25 <- data.frame(Sex = "Male", Age = 25)
m_25_logodds <- predict(donner_blr, newdata = m_25, type = "link")
m_25_logodds
m_25_prob <- predict(donner_blr, newdata = m_25, type = "response")
m_25_prob

```

e. Create new sequence of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.

predicted_probs <- predict(donner_blr, newdata = donner_newdata, type = "response", se.fit = TRUE)

# Coerce outcome into data frame. 

graph_data <- data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)
colnames(graph_data) <- c("Age","Sex","Probability","SE")

```

g. Graph results.

```{r donner_graph}

ggplot(graph_data, aes(x = Age, y = Probability)) +
  geom_line(aes(colour = Sex)) +
  geom_ribbon(aes(ymin = Probability - SE, ymax = Probability + SE, fill = Sex), alpha = 0.3) +
  theme_bw() +
  ylab("Probability of Survival")
  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}
si_full <- list.files(pattern = "solar_irradiation_*") %>% 
  map_df(~read_csv(.)) %>% # Apply read_csv function to all, then put them into a df
  clean_names() # janitor() function to get all column names in snake_case

```

b. Wrangle the data
```{r wrangle_solar}

solar_tidy <- si_full %>% 
  clean_names() %>% 
  rename(sol_rad = etr_wh_m_2, 
         date = yyyy_mm_dd, 
         time = hh_mm_lst) %>% 
  filter(time != "NA") %>% 
  mutate(site = fct_relevel(site, "Hawaii","Santa Barbara","Alaska"))
```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

solar_tidy$date <- mdy(solar_tidy$date) # Now it's a date! 
solar_tidy$time <- hms(solar_tidy$time) # And this is a time! 
```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r}

# NOTE: date_format("%b") could be problematic - start with minimum viable graph option (below)
solar_gg <- ggplot(solar_tidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) +
  scale_fill_gradientn(colors = c("royalblue2","mediumorchid1", "orange", "yellow")) +
  scale_x_date(date_labels ="%b", expand = c(0,0), breaks = '1 month') +
  scale_y_time(breaks = c(21600, 43200,64800), labels = c("6am","12pm","6pm"), expand = c(0,0)) +
  labs(x = "", y = "") +
  theme_classic() +
  theme(axis.line = element_blank(), axis.ticks = element_blank(),axis.text.x = element_text(size = 9, hjust = 0, angle = 45, vjust = 0.5), axis.text.y = element_text(size = 11))  +
  facet_grid(site ~ .)

solar_gg


#### BELOW: MINIMUM VIABLE GRAPH VERSION

solar_gg_min <- ggplot(solar_tidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) +
  scale_fill_gradientn(colors = c("royalblue2","mediumorchid1", "orange", "yellow")) +
  scale_y_time() +
  facet_grid(site ~ .)

solar_gg_min

```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States and Central American countries

a. Get the data, and do some wrangling:
```{r aqua_data}

aq_df <- read_csv("aq_wb.csv")

aq_tidy <- aq_df %>% 
  filter(country == "Brazil" | 
           country == "Chile" | 
           country == "Ecuador" |
           country == "United States") %>% 
  gather(year, aq_prod, '1960':'2014') %>% 
  filter(year >= 1990) %>% 
  mutate(aq_mil = aq_prod/1000000) %>% 
  select(country, year, aq_mil)

```

b. Read in the fish.png as 'fish'
```{r fish_image}

fish <- "fish.png"

```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aq_plot2 <- ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country)) + 
  geom_line(aes(color = country), size = 1) +
  geom_point(aes(color = country)) +
  geom_image(aes(image = fish)) + # Optional
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x = 1), size = 5) +
  scale_color_manual(values = c("darkcyan","goldenrod3","darkmagenta","chocolate1")) + # optional
  labs(x = "Year", y = expression(Aquaculture~Production~(10^6~metric~tons))) + # optional
  coord_cartesian(clip = 'off') + 
  transition_reveal(country, as.numeric(year)) + # How the transitions will occur (lots of options here)
  theme_bw() +
  theme(legend.position = "NA", panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        axis.text.x = element_text(size=12), axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size=14, face = "bold"),
        axis.title.y = element_text(size=14,face="bold")) +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  scale_y_continuous(expand = c(0,0))

aq_plot2

animate(aq_plot2, nframes = 48, renderer = gifski_renderer("aq_animate.gif")) # optional (if they want to save and post/send)

############
# BELOW: MINIMUM VIABLE PLOT ABOVE:
#############

aq_plot3 <- ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country)) + 
  geom_line(aes(color = country), size = 1) +
  geom_point(aes(color = country)) +
  geom_image(aes(image = fish)) + # Optional
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x = 1), size = 5) +
transition_reveal(country, as.numeric(year))  # How the transitions will occur (lots of options here)

aq_plot3

# To render: 
animate(aq_plot3, nframes = 48, renderer = gifski_renderer("aq_animate.gif")) # optional (if they want to save and post/send)

```

##END LAB